---
title: "2. Analyse Health Data"
description: >
  Account lists
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{2. Analyse Health Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include = F,warning = F,message = F}
set.seed(2903)
library(tidyverse)
base_path <- ".."
parsed_healthdata <- 
  readRDS(file.path(base_path,
                    "dev/data/parsed_healthdata_mp.rds")) %>%
  mutate(sourceName = sourceName %>% str_replace_all("Mariana","XXX")) %>%
  select(-sourceVersion,-device,-creationDate)

```

## Analysing Health Data

```{r,message = F,warning = F}
library(health2R)
library(tidyverse)
library(lubridate)
library(purrr)
```
```{r,eval = F}
parsed_healthdata <- read_rds("path/to/your/health_data.rds")
```

The raw `xml` file exported by Apple contains several dozens of variables, which are all exported as rows in a single file. 

```{r}
parsed_healthdata$type %>% table() %>% sort(decreasing = TRUE) %>% head(10)
```

To better make sense and structure the data, this package proposes a number of domain-specific functions to analyse a specific health feature, such as cardiovascular data. 

### Analyse Heart Rate Data

The dedicated function for extracting and analysing cardiovascular data is called `get_cardio` and takes as input an already parsed healthdata object, or the raw zip file. Furthermore, the variable `var` allows the user to pick which cardio-related variable should be extracted, and currently takes as options `heartrate`, `resting`, `walkingAverage`, `workout`, `variability`, and `recovery`.

```{r,eval = T}
hr_data <- get_cardio(health_db = parsed_healthdata, 
                          var = "heartrate")
hr_data %>% select(type,sourceName,unit,startDate,value) %>% head()
```

```{r,warning = F}
library(ggplot2)

ggplot(hr_data, 
       aes(x = startDate, y = value)) + 
    geom_text(aes(label = "\u2665"), color = "red", size = 3) +
    scale_x_datetime(limits = c(as.POSIXct("2022-03-29 00:00:00"),
                                as.POSIXct("2022-03-29 23:59:59")))
```

### Analyse VO2-Max

Furthermore, the `get_vo2max` function obtains and the VO2-Max performance data, either for the full available time range (default), of for a specific `start_date` and `end_date`. 

```{r,eval = T}
vo2max_data <- get_vo2max(health_db = parsed_healthdata,
                          start_date = as_date("2023-01-01"), 
                          end_date = as_date("2023-11-01"))
```

We can now VO2-Max (with differential noise added for privacy reasons):
```{r,include = F}
set.seed(2903)
```

```{r}
private_vo2 <- 
  vo2max_data %>% 
  mutate(value = value + rnorm(1,0,10)) 
private_vo2 %>% head()
```

Plot and smoothed trend overtime: 

```{r}
ggplot(private_vo2, 
       aes(x = startDate, 
           y = value)) + 
  geom_point() +
  scale_x_datetime(limits = c(as.POSIXct("2023-01-29 00:00:00"),
                                as.POSIXct("2023-10-01 00:00:00")),
                     date_breaks = "1 month",
                     date_labels = "%b %y") + 
  geom_smooth() + 
  theme_bw() + 
  theme(axis.text.y = element_blank()) + 
  labs(x = NULL, y = NULL,
       title = "VO2-Max over time (2023)",
       subtitle = "White noise added for additional privacy")
```

### Analyse Menstrual Cycle

```{r}
menstruation_cycle_data <- get_menstruation_cycle(health_db = parsed_healthdata)
```

Of course, female menstrual cycles are a sensitive piece of data, which would not be appropriate to be exposed in this online documentation. Instead of plotting someone's actual cycle data, we therefore simulate random data in a similar format as they would be exported by the `get_menstruation_cycle` function of the package. 

```{r}
set.seed(12345) # for reproducibility

# Function to simulate a single cycle
simulate_cycle <- function(start_date) {
  cycle_length <- sample(25:35, 1) # Random cycle length between 25 and 35 days
  menstruation_length <- sample(3:6, 1) # Random menstruation length between 3 and 7 days
  data.frame(
    Date = seq.Date(start_date, by = "day", length.out = cycle_length),
    Menstruation = c(rep(TRUE, menstruation_length), rep(FALSE, cycle_length - menstruation_length))
  )
}

# Simulate cycles for a year
simulate_year_cycles <- function(start_date, end_date) {
  current_date <- start_date
  all_cycles <- data.frame(Date = as.Date(character()), Menstruation = logical())
  
  while(current_date < end_date) {
    cycle <- simulate_cycle(current_date)
    all_cycles <- bind_rows(all_cycles, cycle)
    current_date <- max(cycle$Date) + 1
  }
  all_cycles %>% filter(Date <= end_date)
}

# Use the function to simulate cycles between two dates
start_date <- as.Date("2024-01-01")
end_date <- as.Date("2024-12-31")
simulated_menstrual_data <- simulate_year_cycles(start_date, end_date)

# Viewing the first 10 rows
simulated_menstrual_data %>% slice(25:35)
```

```{r}
simulated_menstrual_data %>%
  mutate(Month = floor_date(Date, "month"), 
         Day = day(Date)) %>%  
  ggplot(aes(x = Day,  y = 1,fill = Menstruation)) + 
  geom_tile(color = "white") + # Use tiles
  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "grey")) + 
 facet_wrap(~Month, scales = "free_y", 
             labeller = labeller(Month = function(x) format(as.Date(as.character(x), "%Y-%m-%d"), "%b %Y"))) +  # Custom labels
  labs(fill = "Menstruation", x = "Day of the Month", y = "") + 
  theme(axis.text.y = element_blank(), 
        strip.text.y = element_text(angle = 0)) 

```

