---
title: "3. Add Differential Privacy"
description: >
  Account lists
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{3. Add Differential Privacy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include = F}
base_path <- ".."
parsed_healthdata <- 
  readRDS(file.path(base_path,
                    "dev/data/parsed_healthdata_mp.rds"))

```

## Adding Differential Privacy

Health data data can repesent quite sensitive pieces of information, which in some cases should be avoided to be leaked in an inappropriate context. To still be able to use one's wearable data, the `health2R` package provides a function called `add_differential_privacy`, which takes a duration object as input for the variable `max_differential_privacy_offset`, and subsequently makes a data frame differentially private by adding random noise to all time stamps in the data, while retaining the relationship *between* the respective time data. 

```{r,message = F,warning = F}
library(health2R)
library(tidyverse)
library(lubridate)
```

First we plot some example heart rate data:

```{r,eval = F}
parsed_healthdata <- read_rds("path/to/your/health_data.rds")
```
```{r,eval = T}
rhr_data <- get_cardio(health_db = parsed_healthdata, 
                          var = "heartrate",
                       start_date = as_date("2023-02-28"),
                       end_date = as_date("2023-03-02"))
rhr_data %>% 
  select(type,sourceName,unit,startDate,value) %>% 
  head()
```

And the differentially privat data, which has been privatized with a random noise addition of up to 2 days. This duration object parameter can also hold shorter periods as appropriate, e.g. `duration(3600,"seconds")` or similar. 

```{r}
priv_data <- 
  rhr_data %>%
  select(type,sourceName,unit,startDate,value) %>% 
  add_differential_privacy(max_differential_privacy_offset = lubridate::duration(2,"days")) 

priv_data %>% head()
```

Plotting the two time series shows the successful addition of obfuscating noise:

```{r}
combined_data <- bind_rows(rhr_data %>% mutate(type = "baseline_data"), 
                       priv_data %>% mutate(type = "differentially_private_data"))

ggplot(combined_data,
       aes(x = startDate,
           y = value, 
           col = type)) + 
  geom_line() +
  facet_wrap(~type,nrow = 2) + 
  scale_x_datetime(date_breaks = "1 day",
                   date_labels = "%b %d %H:%M UTC")  + 
  theme(axis.text.x = element_text(angle = 45,vjust = 1,hjust = 1))

```

