---
title: "4. Analyse Workout Data"
description: >
  Account lists
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{4. Analyse Workout Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include = F}
base_path <- ".."
parsed_healthdata <- 
  readRDS(file.path(base_path,
                    "dev/data/parsed_healthdata_mp.rds"))

base_path <- ".."
parsed_workouts <- 
  readRDS(file.path(base_path,
                    "dev/data/parsed_apple_workouts_mp.rds"))
```

## Plotting Workouts 

A final function family allows user to analyse and visualise workout data after parsing them from the respective `gpx` format provided by apple. 

```{r,message = F,warning = F}
library(health2R)
library(tidyverse)
library(lubridate)
library(sf)
```

We first load the list of all workouts which we previously parsed in `vignette("01_import_apple_data")` and pick one specific workout:

```{r,eval = F}
parsed_workouts <- 
  readRDS(file.path("path/to/workouts/rds"))
```

```{r}
workout_data <- parsed_workouts$`Route 2023-10-08 12:35pm`

sf_data <- workout_data %>%
    st_as_sf(coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs = 3857) # Using Web Mercator for visualization
```

To not leak any specific protected location, such as one's home, we create a function that removes the first and last `x` minutes of a workout: 

```{r}
remove_start_end <- function(sf_data, minutes) {
  # Convert 'minutes' to seconds since the 'time' column is in POSIXct
  seconds <- minutes * 60
  
  # Get the start and end time
  start_time <- min(sf_data$time)
  end_time <- max(sf_data$time)
  
  # Calculate the time limits
  start_limit <- start_time + seconds
  end_limit <- end_time - seconds
  
  # Filter the data to exclude points within the first and last x minutes
  sf_data_filtered <- sf_data %>%
    filter(time > start_limit & time < end_limit)
  
  return(sf_data_filtered)
}
```

```{r}
sf_data_obscured <- remove_start_end(sf_data, minutes = 5)

```

We can then plot the workout and the respective velocity:

```{r}
ggplot(data = sf_data_obscured) +
  geom_sf(aes(color = speed), size = 1, alpha = 0.7) + 
  scale_color_viridis_c(name = "Speed (m/s)", 
                        option = "C", 
                        limits = range(sf_data$speed, na.rm = TRUE)) +
  theme_minimal() +
  labs(title = "Workout Route with Speed Dynamics",
       x = "Longitude",
       y = "Latitude") +
  theme(legend.position = "right")
```

To plot the workout against a real city map of Berlin, we use the `d6berlin` package: 

```{r}
#devtools::install_github("EcoDynIZW/d6berlin")
library(d6berlin)
```

```{r,warning = F,message = F}
base_map <- base_map_imp(color_intensity = 1, resolution = 500, globe = TRUE)

# Combine the base map with your data
berlin_map <- 
  base_map +
  geom_sf(data = sf_data_obscured, aes(color = speed), col = "red",size = 1, alpha = 0.7) +
  labs(title = "Workout Route with Speed Dynamics")+
   guides(color = FALSE)

# Print the combined map
print(berlin_map)
```
